---
title: "HW 2"
author: Shihui Zhu sz3029
output: github_document
---

```{r package, message=FALSE}
# This chunk loads all the packages used in this homework
library(readxl)
library(tidyverse)
```

# Problem 1

## Read in Mr. Trash Wheel dataset and clean the data
```{r}
mr_trash_wheel_df <- 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = 1,
  range = "A2:N408") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(sports_balls = round(sports_balls, digits = 0))
```

## Read in precipitation data for 2018
```{r}
preci_18_df <- 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = 5,
  range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = "2018")
```

## Read in precipitation data for 2019
```{r}
preci_19_df <- 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = 4,
  range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = "2019")
```

## Combine precipitation data for 2018 and 2019, and convert month to character variable
```{r}
preci_18_and_19_df <- 
  bind_rows(preci_18_df, preci_19_df) %>%
  mutate(month = month.name[month])
```

## Description for Mr. Trash Wheel dataset

The Mr. Trash Wheel dataset contains `r length(names(mr_trash_wheel_df))` variables:
`r names(mr_trash_wheel_df)`. It contains `r nrow(mr_trash_wheel_df)` rows 
and `r ncol(mr_trash_wheel_df)` columns of observations. The data were 
collected from years `r unique(dplyr::pull(mr_trash_wheel_df, year))`. The total weight 
of trash collected over the 6 years was `r sum(dplyr::pull(mr_trash_wheel_df, weight_tons))` 
tons, with a total volume of `r sum(dplyr::pull(mr_trash_wheel_df, volume_cubic_yards))` 
cubic yards. A table displaying the values of the key variables is shown below: 

`r mr_trash_wheel_df %>% dplyr::group_by(year) %>% select(year, weight_tons:homes_powered) %>% skimr::skim() %>% knitr::kable()`

There were different kinds of trash collected. For example, the total number of sports balls collected in 2019
was `r mr_trash_wheel_df %>% dplyr::filter(str_detect(year, "2019")) %>% dplyr::pull(sports_balls) %>% sum()`, and the median number is `r mr_trash_wheel_df %>% dplyr::filter(str_detect(year, "2019")) %>% dplyr::pull(sports_balls) %>% median()`. 

The precipitation dataset contains total precipitation per month from 2018 to 2019. It has `r length(names(preci_18_and_19_df))` variables: `r names(preci_18_and_19_df)`, with a dimension of `r nrow(preci_18_and_19_df)` rows and `r ncol(preci_18_and_19_df)` columns. The total precipitation in year 2018 was `r preci_18_and_19_df %>% dplyr::filter(str_detect(year, "2018")) %>% dplyr::pull(total) %>% sum()`. 

## Description for precipitation data for 2018 and 2019





# Problem 2
