---
title: "HW 2"
author: Shihui Zhu sz3029
output: github_document
---

```{r package, message=FALSE}
# This chunk loads all the packages used in this homework
library(readxl)
library(tidyverse)
```

# Problem 1

## Read in Mr. Trash Wheel dataset and clean the data
```{r trash_wheel}
mr_trash_wheel_df <- 
  read_excel("data/p1/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = 1,
  range = "A2:N534") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(sports_balls = round(sports_balls, digits = 0)) # Round the number of sports balls to nearest integer
```

## Read in precipitation data for 2018
```{r preci18}
preci_18_df <- 
  read_excel("data/p1/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = 7,
  range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = "2018") # Add year variable
```

## Read in precipitation data for 2019
```{r perci19}
preci_19_df <- 
  read_excel("data/p1/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = 6,
  range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = "2019") # Add year variable
```

## Combine precipitation data for 2018 and 2019, and convert month to character variable
```{r combine1819}
preci_18_and_19_df <- 
  bind_rows(preci_18_df, preci_19_df) %>%
  mutate(month = month.name[month]) # convert month to character variable

preci_18_and_19_df
```

## Description for Mr. Trash Wheel dataset

The Mr. Trash Wheel dataset contains `r nrow(mr_trash_wheel_df)` observations 
of `r ncol(mr_trash_wheel_df)` variables. The variables are:
`r names(mr_trash_wheel_df)`. Columns from `plastic_bottles` to `homes_powered` indicate
different kinds of trash collected at Mr. Trash Wheel. The data were collected from years `r unique(dplyr::pull(mr_trash_wheel_df, year))`.

The median number of sports balls collected in 2019 was `r mr_trash_wheel_df %>% dplyr::filter(year == "2019") %>% dplyr::pull(sports_balls) %>% median()`. 

## Description for precipitation data for 2018 and 2019

The precipitation dataset contains data of total precipitation per month from 2018 to 2019. It has `r length(names(preci_18_and_19_df))` variables: `r names(preci_18_and_19_df)`, and total `r nrow(preci_18_and_19_df)` observations. The total precipitation in year 2018 was `r preci_18_and_19_df %>% dplyr::filter(year == "2018") %>% dplyr::pull(total) %>% sum()`.


# Problem 2
## Clean data in pols-month.csv
* break up the variable `mon` into integer variables `year`, `month`, and `day`
* replace month number with month name
* create a `president` variable taking values `gop` and `dem`, and remove `prez_dem` and `prez_gop`
* remove the `day` variable
```{r pols, message=FALSE}
pols <- read_csv("data/p2/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), convert = TRUE) %>% # convert the format of date
  mutate(month = month.name[month], # convert month name to string
         president = recode(prez_dem, `1` = "dem", `0` = "gop")) %>% # create president variable and assign value
  select(-day, -prez_dem, -prez_gop) # remove unnecessary columns

pols
```

## Clean data in snp.csv
Use similar process as above:

* break up the variable `date` into integer variables `year`, `month`, and `day`
* replace month number with month name
* remove `day` variable
```{r snp, message=FALSE}
snp <- read_csv("data/p2/snp.csv") %>%
  janitor::clean_names() %>%
  mutate(date = as.Date(as.character(date),"%m/%d/%y"),
         date = as.Date(ifelse(date > Sys.Date(),
                        format(date, "19%y-%m-%d"),
                        format(date)))) %>% 
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>% # reformat the date
  select(year, month, everything()) %>% # make year and month as the leading columns
  mutate(month = month.name[month]) %>% # convert month name to string
  select(-day) # remove day column

snp
```

## Tidy unemployment data

* switching from “wide” to “long” format
* match the variable names

```{r unemploy, message=FALSE}
unemployment <- read_csv("data/p2/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_percent"
  ) %>%
  mutate(month = match(month, month.abb),
         month = month.name[month]) %>% # reformat the date
  janitor::clean_names()

unemployment
```
## Merge the datasets

* merge `snp` into `pols` via `left_join()`
* merge with `unemployment` data
```{r merge}
pols_snp <-
  left_join(pols, snp, by = c("year", "month"))
pols_snp_unemployment <-
  left_join(pols_snp, unemployment, by = c("year", "month"))

pols_snp_unemployment
```

## Descriptions for datasets: `pols`, `snp`, `unemployment`, and `pols_snp_unemployment` merged from those

1. The `pols` dataset contains `r nrow(pols)` observations from `r length(names(pols))` variables.
The variables are `r names(pols)`. This dataset contains information related to 
the number of national politicians who are democratic or republican at any given time, ranging from `r min(dplyr::pull(pols, year))` to `r max(dplyr::pull(pols, year))` i.e. range of `r max(dplyr::pull(pols, year)) - min(dplyr::pull(pols, year))` years. 

2. The `snp` dataset contains `r nrow(snp)` observations from `r length(names(snp))` variables.
The variables are `r names(snp)`. This dataset contains information related to 
Standard & Poor’s stock market index (S&P), often used as a representative measure 
of stock market as a whole, ranging from `r min(dplyr::pull(snp, year))` 
to `r max(dplyr::pull(snp, year))` i.e. 
range of `r max(dplyr::pull(snp, year)) - min(dplyr::pull(snp, year))` years.

3. The `unemployment` dataset contains `r nrow(unemployment)` observations from `r length(names(unemployment))` variables. The variables are `r names(unemployment)`. This dataset contains information about percentage of unemployment at any given time, ranging from `r min(dplyr::pull(unemployment, year))` to `r max(dplyr::pull(unemployment, year))` i.e. range of `r max(dplyr::pull(unemployment, year)) - min(dplyr::pull(unemployment, year))` years. 

4. The merged dataset from the above three, `pols_snp_unemployment`, contains `r nrow(pols_snp_unemployment)` observations from `r length(names(pols_snp_unemployment))` variables. The variables are `r names(pols_snp_unemployment)`. This dataset contains information related to the number of national politicians who are democratic or republican, the Standard & Poor’s stock market index (S&P), and the percentage of unemployment at any given time, ranging from `r min(dplyr::pull(pols_snp_unemployment, year))` to `r max(dplyr::pull(pols_snp_unemployment, year))` i.e. range of `r max(dplyr::pull(pols_snp_unemployment, year)) - min(dplyr::pull(pols_snp_unemployment, year))` years.

# Problem 3
## Load and tidy the dataset for `Popular_Baby_Names.csv`

* convert all characters to upper cases, make the labels uniform
* remove duplicated rows
```{r baby, message=FALSE}
baby_name <- read_csv("data/p3/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(gender = str_to_upper(gender),
         ethnicity = str_to_upper(ethnicity),
         childs_first_name = str_to_upper(childs_first_name),
         ethnicity = replace(ethnicity, ethnicity == "ASIAN AND PACI", "ASIAN AND PACIFIC ISLANDER"),
         ethnicity = replace(ethnicity, ethnicity == "BLACK NON HISP", "BLACK NON HISPANIC"),
         ethnicity = replace(ethnicity, ethnicity == "WHITE NON HISP", "WHITE NON HISPANIC")) %>% 
  distinct() # remove duplicated rows

baby_name
```


## Create a table of rank in popularity of the name “Olivia” as a female baby name over time

```{r olivia}
olivia <-
  baby_name %>%
  dplyr::filter(childs_first_name == "OLIVIA") %>%
  select(-childs_first_name, -gender) %>%
  arrange(year_of_birth) %>%
  knitr::kable(
    col.names = c("Year of Birth", "Ethnicity", "Count", "Rank"),
    caption = "The Table of Rank in Popularity of the Name “Olivia” As a Female Baby Name Over Time"
  )
olivia
```
## Create a table of the most popular name among male children over time

```{r male}
male <-
  baby_name %>%
  dplyr::filter(gender == "MALE",
                rank == '1') %>%
  select(-gender, -rank) %>%
  arrange(year_of_birth) %>%
  knitr::kable(
    col.names = c("Year of Birth", "Ethnicity", "Children's Names", "Count"),
    caption = "The Table of the Most Popular Name Among Male Children Over Time"
  )
male
```

## Scatter Plot

* For male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis)
```{r plot}
# Select the names
male_nh <-
  baby_name %>%
  dplyr::filter(gender == "MALE",
                year_of_birth == '2016',
                ethnicity == "WHITE NON HISPANIC") %>%
  select(rank, childs_first_name, count)

# Create the plot
ggplot(male_nh, aes(x = rank, y = count)) +
  geom_point() +
  xlab("Rank") +
  ylab("Number") +
  ggtitle("Number v.s. Rank for male, white non-hispanic children born in 2016")

## Save the plot
ggsave("hw2_p2_scatter_plot.jpg")
```



